One-Step, Three-Factor Authentication via an Earpiece

### 2017-04-01 paper.tex
John's changes to related work, overall reducing text

\title{One-Step, Three-Factor Authentication via Ear EEG}

\numberofauthors{4}
\author{
  \alignauthor 1st Author\\
    \affaddr{Affiliation}\\
    \email{e-mail}\\
  \alignauthor 2nd Author\\
    \affaddr{Affiliation}\\
    \email{e-mail}\\
  \alignauthor 3rd Author\\
    \affaddr{Affiliation}\\
    \email{e-mail}\\
  \alignauthor 4th Author\\
    \affaddr{Affiliation}\\
    \email{e-mail}\\
}

@@

\keywords{
  usable security; multi-factor authentication; wearable authentication; bio-sensory computing; passthoughts
}

@@

\subsection{Passthoughts Authentication}
The use of EEG as a biometric signal for user authentication has a relatively short history.
In 2005, Thorpe et al. motivated and outlined the design of a passthoughts system \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-
100\% authentication accuracy using multi-channel sensors placed on the scalp \cite{Poulos2002,Marcel2007a,Palaniappan2008,Ashby2011}.
In 2013, one group showed that 99\% authentication accuracy can also be
achieved using a consumer-grade single-channel sensor on the scalp \cite{Chuang2013b}. In particular, they proposed that the
lack of signal diversity from multiple EEG channels can be overcome by allowing users to choose their own personalized passthoughts (e.g., sing their favorite song in their head). There are two significant consequences of this result. First,
the passthoughts approach is no longer constrained by the high cost (\textgreater \$10,000 USD)
and low usability (gel-based electrodes in hair; aesthetic challenges of a full EEG cap) of
medical-grade multi-channel devices. Second, because users can choose and
easily change their secret mental task, this approach can support one-step two-
factor authentication via the simultaneous presentation of the inherence factor
(brainwave signatures due to the unique folding structures of the cortex) and the
knowledge factor (the secret mental task) \cite{Chuang2014}.

By employing scalp-based consumer-grade electroencephalography (EEG), it was demonstrated in a 2014 passthoughts study that a user can submit both a knowledge factor (i.e., secret thought) and an inherence factor (i.e., brainwave signal unique to the individual) in a single step by performing a single mental task \cite{Chuang2014}. Additionally, the robustness of this method against impersonation attacks was demonstrated, including conditions where the attacker may have learned the target's secret thought and/or secret task \cite{Johnson2014}.

Following this, we investigate custom-fit in-ear EEG technology as the platform for investigating the feasibility, performance, and usability of one-step three-factor authentication worn unobtrusively. The system we propose and test here uses the choice of a mental task or "passthought" to perform as knowledge (factor one), the uniqueness of an individual's brain activity as measured by EEG as inherence (factor two), and the physical token of a custom-fit earpiece that could easily contain a hardware key-pair, as a possession factor (factor 3). Because three-factor authentication (3FA) requires the user to submit one distinct instance of each type of authenticator, it represents the strongest level of authentication security possible. The ability to utilize all three of these security factors in a single step by performing a mental task of a few seconds is promising in pursuit of extremely strong security while maintaining a low amount of effort and obtrusiveness to the user, and could even be enjoyable.

\subsection{In-Ear EEG}
Even consumer-grade headsets can be uncomfortable to wear, and are awkwardly visible to outside observers. Sensors placed around the ear, however, present a far more discreet, comfortable location for EEG electrodes, as many people already wear technology like earbuds in day-to-day life.

Research in in-ear EEG is only several years old. Nonetheless, the concept has
attracted attention because of the discreetness factor of in-ear EEG over
traditional scalp-based EEG. A research team at the Imperial College London
and Aarhus University published a landmark paper in 2011 that introduced the
concept of in-ear EEG, demonstrating for the first time the feasibility of recording
brainwave signals from within the ear canal \cite{Looney2011}.
Follow-up work from the same group demonstrated its ability to produce signal-to-noise ratios comparable to
those from conventional EEG electrode placements, robustness to common
sources of artifacts, and use in a brain-computer interface (BCI) system based on
auditory evoked potentials and visual evoked potentials
\cite{Looney2012a,Kidmose2013a,Kidmose2013b}.
United Sciences is currently developing a consumer "hearable" (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics \cite{UnitedSciences}.

\cite{curran2016passthoughts} was the first to merge in-ear EEG with passthought authentication,
 using a minimally modified consumer-grade EEG device with a single electrode, achieving approximately 80\% authentication accuracy.

\subsection{One-Step, Multi-Factor Authentication}
Behavioral authentication methods such as keystroke dynamics and speaker authentication can be categorized as one-step two-factor authentication schemes. In both cases, the knowledge factor (password or passphrase) and inherence factor (typing rhythm or speaker's voice) are employed \cite{Monrose1997}. In contrast, the Nymi band supports one-step two-factor authentication via the inherence factor (cardiac rhythm that is supposed to be unique to each individual) and the possession factor (the wearing of the band on the wrist) \cite{Nymi}. However, as far as we know, no one has proposed or demonstrated a one-step three-factor authentication scheme, in which possession of a unique device also serves to authenticate the user. In this paper, we introduce custom-built EEG devices, incorporating an added possession factor to the already one-step two-factor authentication of passthoughts.

@@

When proposing or evaluating authentication paradigms, robustness against imposters is often the first consideration, but the usability of these systems is of equal importantance as they must conform to a person's needs and lifestyle to warrant adoption and prolonged use. In \cite{sasse2001}, the authors describe usability issues with knowledge-based systems like alphanumeric passwords, in particular that it should not be the burden of the user to remember complex passwords that have to be frequently changed. \cite{braz2006} analyze some of the complexities applying traditional interface human factors heuristics to authentication, and indicate the importance of social acceptibility, learnability, and simplicity of authentication methods.

Head-worn devices face particular usability issues; in their analysis of user perceptions of such devices, \cite{Genaro2014} identify design, usability, ease of use, and obtrusiveness among the top 10 concerns of users of these devices and qualitative concerns about comfort or "looking weird".

Mobile and wearable technologies' continuous proximity to the user's body provides favorable conditions for unobtrusively capturing biometrics for authentication. Many such uses have been proposed that embrace usability such as touch-based interactions \cite{Tartz2015, Holz2015} and walking patterns \cite{Lu2014} using mobile phones, as well as identification via head movements and blinking in head-worn devices \cite{Rogers2015}. However, these typically draw only from the inherence factor (though could incorporate the possession factor of the device). \cite{Chen2015} proposed an inherence and knowledge two-factor method for multi-touch mobile devices based on a user's unique finger tapping of a song, though this may be vulnerable to "shoulder surfing": imposters observing and mimicking the behavior. By comparison, passthoughts are largely invisible to attackers. This system also presents a solution to "rubber-hose attacks" (forceful coercion to divulge a password): a thought, which is secret (and thus changeable), but has a particular expression unique to an individual, the performance of which cannot be described (and thus cannot be coerced).

%% 3/28

@@

Seven male, right-handed participants (P1-P7), 5 students and 2 non-students, completed our study protocol approved by our local ethics review board. Though this sample is relatively homogenous, this quality interestingly functions to strengthen the results of this initial exploration (see details in Discussion). After participants' 3D ear molds were obtained and their fit and electrical impedance was checked, the collection of data for authentication analyses commenced. Data collection consisted of participants completing a short questionnaire, a set up period with the OpenBCI system and earpieces including a second impedance check, the performance of 9 mental tasks presented on a laptop, and finally a post-experiment questionnaire.

@@

To produce custom ear impressions during the first study visit, we cleaned subjects' ears and injected silicon into the ear canals. A cotton ball with a string attached is placed into the ear canal first.  When the silicon dries after a few minutes, the string is pulled to remove the impression from the ear canal. This impression is then scanned with a 3D scanner and the resulting scan modified based on some heuristics to achieve a comfortable fit and to ensure the intended electrode sites will make good contact with the skin. Channels are created in the 3D model to allow wire leads and associated EEG electrodes as well as a plastic tube to deliver audio. This 3D model is then sent to a 3D printer. Following this wires, leads, and associated AgCl electrodes are installed.

The positions of the earpiece electrodes were simplified from those described in \cite{Mikkelsen2015}. We reduced the number of canal electrodes in order to prevent electrical bridging and positioned them approximately 180 degrees apart in the canal (posterior/back and anterior/front locations in the canal). We also reduced the number of electrodes in the concha to one. An example of one of the manufactured earpieces is shown in Figure \ref{fig:earpiece_diagram}.

@@

\subsection{Data Collection Protocol}
%\label{sec:org2041857}
All sites were cleaned with ethanol prior to electrode placement and a small amount of conductive gel was used on each electrode. For EEG recording we used OpenBCI \cite{michalska2009openbci}, an open-source biosensing system to maintain an overall low cost for our setup. OpenBCI costs about 600 USD, and thus is an affordable alternative to medical-grade EEG systems (about 20,000 USD), and has demonstrated effectiveness for non-medical use cases such as this one \cite{Frey2016}. The OpenBCI system we used allows for 8 channels of simultaneous recording, along with separate ground and reference channels. Data was collected with the ground placed at the center of the forehead, approximately AFz according to the 10-20 International Standard for Electrode Placement (ISEP), and using the left mastoid as reference. 6 channels were used for the three electrodes on Eech earpiece (shown in Figure \ref{fig:earpiece_diagram}), for the remaining two channels AgCl ring electrodes were placed on the right mastoid for later re-referencing, and at approximately Fp1 (above the left eye in in the ISEP) for validating the data collected in the ears against a common scalp-based placement. Before beginning the experiment, the data from each channel was visually inspected using the OpenBCI interface and participants were asked to blink and clench their jaws to confirm that all channels were active and properly connected. Audio stimuli were delivered through small tubes in the earpieces.

During the experiment, participants were seated in a comfortable position in a quiet room facing a laptop on which the instructions and stimuli were presented using PsychoPy \cite{peirce2007psychopy}. The entire set of tasks was completed once over with five trials each, followed by another five trials each in order to reduce boredom and repetition effects of a single task. Each trial was 10 seconds in length, for a total of 10 trials and 100 seconds of data collected per task. The instructions were read aloud to the participant by the experimenter, and the experiment was advanced using a pointer held in the participant's lap to minimize motion artifacts in the data. The experimenter also recorded the participant's chosen secrets for the \textit{sport}, \textit{song}, \textit{face}, \textit{speech}, and \textit{sequence} tasks and reminded the participant of these for the second set of trials.

 After completing the experiment, a subset of participants completed a usability questionnaire in which they were asked to rate their ease of performing, level of engagement, perceived repeatability, and likeliness of using each task. The questionnaire also asked participants to rank the tasks overall from most to least favorite, as well as several open response questions regarding potential use cases of in-ear EEG and this method of authentication, level of comfort wearing the earpieces, and any other comments they chose to provide.

@@

We were able to confirm the custom-fit earpieces are able to collect EEG data via two metrics: good impedances measured for the ear electrodes, and alpha-band EEG activity attenuation when a participant's eyes were open versus closed.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{lrrrrrr}
& \multicolumn{6}{c}{\textbf{Impedances} [k\(\Omega\)]} \\
\cline{2-7}
& \multicolumn{3}{|c|}{\textbf{Left ear}} & \multicolumn{3}{c|}{\textbf{Right ear}} \\
\textbf{P} & \textbf{C} & \textbf{F} & \textbf{B} & \textbf{C} & \textbf{F} & \textbf{B} \\
\hline
1 & 4 & 4 & 4 & \textless1 & 4 & 3\\
2 & 9 & 5 & 4 & 3 & 4 & 4\\
3 & 4 & 5 & 4 & 9 & 6 & 9\\
4 & 4 & 5 & 4 & 3 & 16 & 9\\
5 & 9 & 20 & 7 & 3 & 7 & 9\\
6 & 5 & 8 & 2 & 1 & 1 & 9\\
7 & 2 & 9 & 8 & 7 & 5 & 6\\
\end{tabular}
\end{center}
\caption{Electrical impedances measured for earpiece electrodes, for concha (C), front (F) and back (B).}
\label{tab:impedances}
\end{table}

For the alpha-attenuation test, data from the \textit{breathe} task was compared with that of the \textit{breathe - open} task. It is a well-known feature of EEG data that activity in the alpha-band (approximately 8-12 Hz range) increases when the eyes are closed compared with a similar state with eyes open. For our participants, this attenuation is clearly visible even in just a single trial's data. We also performed this comparison on the data collected from the Fp1 electrode and see the effect clearly there as well to compare. Figure \ref{fig:alpha_atten} shows the alpha attenuation in the left ear channels, as well as Fp1 of one participant as an example. We see the same effect in the right ear channels.

@@

We analyzed the EEG signals collected during the tasks using a support vector classifier (SVC). Since past work has shown that classification tasks in EEG-based BCI are linear \cite{Garrett2003a}, we used XGBoost, a popular tool for logistic linear classification \cite{Chen2016}. Compared to other linear classifiers, XGBoost uses gradient boosting; in which an algorithm generates an ensemble (in this case, a decision tree) of weak linear classifiers that minimizes a given loss function. Gradient boosting gives better results in linear classification problems, without our needing to manually tune classifier hyperparameter.

@@

For each configuration of electrodes, we calculated the mean FAR and FRR across all participants using each task as the passthought (Figure \ref{fig:meanByElectrode}).
Incorporating all electrodes data results in a perfect score for all tasks.
Using data from all left-ear electrodes achieves the next lowest FAR.
Interestingly, no single electrode performs as well as the aggregate left or both-ear plus Fp1 (all) conditions.

@@

Our results indicate acceptable accuracy using electrodes on the left ear alone. 
This corresponds to our original scenario, in which the device could be worn like an earbud (ideally, only one earbud would need sensors). As such, we focus on results from the left ear alone in our following analyses.

@@

A subset of four participants completed our usability questionnaire. This questionnaire asked them to rank each of the mental tasks on 7-point likert-type scales on ease of use, level of engagement, perceived repeatability, and likeliness to use in a real-world authentication setting. The \textit{breathe} and \textit{listen} (\(\mu\)=6.75) tasks were rated easiest to use. \textit{Sequence} (\(\mu\)=5) and \textit{song} were rated highest in engagement, \textit{Breathe} (\(\mu\)=7) and \textit{listen} (\(\mu\)=6.75) were ranted highest in repeatability. In terms of likeliness of use for authentication, \textit{song-open} (\(\mu\)=5) and \textit{sequence} (\(\mu\)=4.25) were rated highest, though modestly. We also asked participants to rank the tasks from most (1) to least (9) favorite overall. The \textit{song - open} task ranked highest among favorites (\(\mu\)=4.25) followed by a tie between \textit{breathe - open}, \textit{song}, and \textit{speech} (\(\mu\)=4.75). \textit{Sequence} (\(\mu\)=7.75) and \textit{face} (\(\mu\)=6.75) were rated as least favorites.

@@

While our authentication analysis establishes that passthoughts achieve low FAR and FRR when tested against other participants' passthoughts, we do not know how robust passthoughts are against a spoofing attack, in which both a participant's custom-fit electrode, and details of that participant's chosen passthought, are leaked and used by an imposter.

The first aspect of this scenario we tested was the ability of an imposter to wear an earpiece acquired from someone else and achieve viable impedance values for EEG collection based on the fit of the pieces in their ears. P1 tried on each of the other participants' earpieces, which were able to at least physically fit in P1's ears. The impedances of P1 wearing other participants' earpieces were then recorded and are listed in Table \ref{tab:imposter_impedances} below. While some are a better fit than others, overall they are higher than those achieved by the pieces' intended owners themselves.

@@

To explore the scenario of an imposter attempting to gain access, we chose the most vulnerable case participant, P6, whose earpieces P1 had the lowest impedances while wearing. We collected data using the same data collection protocol, but had P1 refer to P6's report of chosen passsthoughts.
P1 performed each of P6's passthoughts (simulating an "inside imposter"). Following the same classification and preprocessing steps, we generated 200 samples per task for our imposters, using data from all left ear electrodes.

@@

Since every participant has one classifier per task (for which that task is the passthought), we are able to make 200 spoofed attempts with the correct passthought on each of P6's classifiers. We find no successful spoof attempts for tasks with a chosen secret (e.g., \textit{song} or \textit{face}). However, we also do not find any successful spoof attacks for tasks with no chosen secret (e.g., \textit{breathe}). In fact, in all 1,800 spoof attempts (200 attempts for each of the 9 classifiers), we do not find a single successful attack on any of P6's classifiers.

However, since this participant's data appeared in the initial pool, the classifier may have been trained on his or her recordings as negative examples. To explore the efficacy of an outsider spoofing recordings, we repeated the same protocol with an individual who did not appear in our initial set of participants (an "outside imposter"). Again, we find zero successful authentication attempts out of 1,800 attempts.

@@

Our findings demonstrate the apparent feasibility of a single earpiece, achieving good results with only three electrodes and a reference, all on or around the left ear. FARs and FRRs are very low across all participants and tasks, with FARs overall lower than FRRs, a desirable pattern in terms of authenticating access to potentially sensitive information. Participants' best-performing passthoughts typically see no errors in our training. Furthermore, no spoofed attacks were successful in our cursory analysis.

>>>>

\title{One-Step, Three-Factor Authentication via an Earpiece}

\numberofauthors{2}
\author{
  \alignauthor 1st Author, 2nd Author, 3rd Author\\
    \affaddr{Affiliation}\\
    \email{1st e-mail, 2nd e-mail, 3rd e-mail}\\
  \alignauthor 4th Author\\
    \affaddr{Affiliation}\\
    \email{e-mail}\\
}

@@

\keywords{
  usable security; multi-factor authentication; wearable authentication; passthoughts; biosensing
}

@@

\subsection{Passthoughts and Behavioral Authentication}
The use of EEG as a biometric signal for user authentication has a relatively short history.
In 2005, Thorpe et al. motivated and outlined the design of a passthoughts system \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-100\% authentication accuracy for small populations using research-grade and consumer-grade scalp-based EEG systems \cite{Poulos2002,Marcel2007a,Ashby2011,Chuang2013b}.
%Since 2002, a number of independent groups have achieved 99-
%100\% authentication accuracy using multi-channel sensors placed on the scalp \cite{Poulos2002,Marcel2007a,Palaniappan2008,Ashby2011}.
%In 2013, one group showed that 99\% authentication accuracy can also be
%achieved using a consumer-grade single-channel sensor on the scalp \cite{Chuang2013b}. In particular, they proposed that the
%lack of signal diversity from multiple EEG channels can be overcome by allowing users to choose their own personalized passthoughts (e.g., sing their favorite song in their head). There are two significant consequences of this result. First,
%the passthoughts approach is no longer constrained by the high cost (\textgreater \$10,000 USD)
%and low usability (gel-based electrodes in hair; aesthetic challenges of a full EEG cap) of
%medical-grade multi-channel devices. Second, because users can choose and
%easily change their secret mental task, this approach can support one-step two-
%factor authentication via the simultaneous presentation of the inherence factor
%(brainwave signatures due to the unique folding structures of the cortex) and the
%knowledge factor (the secret mental task) \cite{Chuang2014}.

%By employing scalp-based consumer-grade electroencephalography (EEG), it was demonstrated in a 2014 passthoughts study that a user can submit both a knowledge factor (i.e., secret thought) and an inherence factor (i.e., brainwave signal unique to the individual) in a single step by performing a single mental task \cite{Chuang2014}. Additionally, the robustness of this method against impersonation attacks was demonstrated, including conditions where the attacker may have learned the target's secret thought and/or secret task \cite{Johnson2014}.

%Following this, we investigate custom-fit in-ear EEG technology as the platform for investigating the feasibility, performance, and usability of one-step three-factor authentication worn unobtrusively. The system we propose and test here uses the choice of a mental task or "passthought" to perform as knowledge (factor one), the uniqueness of an individual's brain activity as measured by EEG as inherence (factor two), and the physical token of a custom-fit earpiece that could easily contain a hardware key-pair, as a possession factor (factor 3). Because three-factor authentication (3FA) requires the user to submit one distinct instance of each type of authenticator, it represents the strongest level of authentication security possible. The ability to utilize all three of these security factors in a single step by performing a mental task of a few seconds is promising in pursuit of extremely strong security while maintaining a low amount of effort and obtrusiveness to the user, and could even be enjoyable.

%\subsection{In-Ear EEG}
%Even consumer-grade headsets can be uncomfortable to wear, and are awkwardly visible to outside observers. Sensors placed around the ear, however, present a far more discreet, comfortable location for EEG electrodes, as many people already wear technology like earbuds in day-to-day life.

The concept of in-ear EEG was introduced in 2011 with a demonstration of the feasibility of recording brainwave signals from within the ear canal \cite{Looney2011}. The in-ear placement can produce signal-to-noise ratios comparable to those from conventional EEG electrode placements, is robust to common sources of artifacts, and can be used in a brain-computer interface (BCI) system based on auditory and visual evoked potentials \cite{Kidmose2013a}. An 80\% accuracy level was achieved for user authentication using in-ear EEG captured with a modified single-channel consumer-grade device \cite{curran2016passthoughts}.

%Research in in-ear EEG is only several years old. Nonetheless, the concept has
%attracted attention because of the discreetness factor of in-ear EEG over
%traditional scalp-based EEG. A research team at the Imperial College London
%and Aarhus University published a landmark paper in 2011 that introduced the
%concept of in-ear EEG, demonstrating for the first time the feasibility of recording
%brainwave signals from within the ear canal \cite{Looney2011}.
%Follow-up work from the same group demonstrated its ability to produce signal-to-noise ratios comparable to
%those from conventional EEG electrode placements, robustness to common
%sources of artifacts, and use in a brain-computer interface (BCI) system based on
%auditory evoked potentials and visual evoked potentials
%\cite{Looney2012a,Kidmose2013a,Kidmose2013b}.
%United Sciences is currently developing a consumer "hearable" (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics \cite{UnitedSciences}.

%\cite{curran2016passthoughts} was the first to merge in-ear EEG with passthought authentication,
 %using a minimally modified consumer-grade EEG device with a single electrode, achieving approximately 80\% authentication accuracy.

%\subsection{One-Step, Multi-Factor Authentication}
Behavioral authentication methods such as keystroke dynamics and speaker authentication can be categorized as one-step two-factor authentication schemes. In both cases, the knowledge factor (password or passphrase) and inherence factor (typing rhythm or speaker's voice) are employed \cite{Monrose1997}. In contrast, the Nymi band supports one-step two-factor authentication via the inherence factor (cardiac rhythm that is supposed to be unique to each individual) and the possession factor (the wearing of the band on the wrist) \cite{Nymi}. However, as far as we know, no one has proposed or demonstrated a one-step three-factor authentication scheme.%, in which possession of a unique device also serves to authenticate the user. %In this paper, we introduce custom-built EEG devices, incorporating an added possession factor to the already one-step two-factor authentication of passthoughts.

@@

When proposing or evaluating authentication paradigms robustness against imposters is often the first consideration, but the usability of these systems is of equal importance as they must conform to a person's needs and lifestyle to warrant adoption and prolonged use. The authors of \cite{sasse2001} describe usability issues with knowledge-based systems like alphanumeric passwords, in particular that it should not be the burden of the user to remember complex passwords that have to be frequently changed. \cite{braz2006} analyzed some of the complexities applying interface human factors heuristics to authentication, and indicate the importance of social acceptibility, learnability, and simplicity of authentication methods. Technology worn on the head faces particular usability issues; in their analysis of user perceptions of such devices, \cite{Genaro2014} identified design, usability, ease of use, and obtrusiveness among the top 10 concerns of users posting online about these devices, as well as qualitative concerns around comfort and "looking weird".

In the proposed system we seek to incorporate the perspectives of this research. Passthoughts such as a favorite song or relaxed breathing are easy for a user to remember and perform, and the choice of an earpiece form-factor of our system seeks to address wearability and social perception concerns.

Mobile and wearable technologies' continuous proximity to the user's body provides favorable conditions for unobtrusively capturing biometrics for authentication. Many such uses have been proposed that embrace usability such as touch-based interactions \cite{Tartz2015, Holz2015} and walking patterns \cite{Lu2014} using mobile phones, as well as identification via head movements and blinking in head-worn devices \cite{Rogers2015}. However, these typically draw only from the inherence factor. \cite{Chen2015} proposed an inherence and knowledge two-factor method for multi-touch mobile devices based on a user's unique finger tapping of a song, though it may be vulnerable to "shoulder surfing": imposters observing and mimicking the behavior. By comparison, passthoughts are largely invisible to attackers. Furthermore, this system also presents a solution to "rubber-hose attacks" (forceful coercion to divulge a password): a thought, which is secret (and thus changeable), but has a particular expression unique to an individual, the performance of which cannot be described (and thus cannot be coerced).

@@

Seven male, right-handed participants (P1-P7), 5 students and 2 non-students, completed our study protocol approved by our local ethics review board. Though this sample is relatively homogenous and greater diversity is necessary for a larger real-world feasibility assessment, this quality interestingly functions to strengthen the results of this initial exploration (see details in Discussion). After participants' 3D ear molds were obtained and their fit and electrical impedance was checked, the collection of data for authentication analyses commenced. Data collection consisted of participants completing a short questionnaire, a set up period with the OpenBCI system and earpieces including a second impedance check, the performance of 9 mental tasks presented on a laptop, and finally a post-experiment questionnaire.

@@

To produce custom ear impressions we first cleaned subjects' ears and injected silicon into the ear canals. A cotton ball with a string attached was placed into the ear canal first. When the silicon dried after a few minutes, the string was pulled to remove the impression from the ear canal. This impression was then scanned with a 3D scanner and the resulting scan modified based on some heuristics to achieve a comfortable fit and to ensure the intended electrode sites will make good contact with the skin. Channels were created in the 3D model to allow wire leads and associated EEG electrodes as well as a plastic tube to deliver audio. This 3D model was then sent to a 3D printer after which wires, leads, and associated AgCl electrodes were installed. The positions of the earpiece electrodes were simplified from those described in \cite{Mikkelsen2015}. We reduced the number of canal electrodes in order to prevent electrical bridging and positioned them approximately 180 degrees apart in the canal (posterior/back and anterior/front locations in the canal). We also reduced the number of electrodes in the concha to one. An example of one of the manufactured earpieces is shown in Figure \ref{fig:earpiece_diagram}.

@@

\subsection{Data Collection Protocol}
%\label{sec:org2041857}
All sites were cleaned with ethanol prior to electrode placement and a small amount of conductive gel was used on each electrode. For EEG recording we used an 8-channel OpenBCI system \cite{michalska2009openbci} which is open-source and costs about 600 USD; an alternative to medical-grade EEG systems (\textgreater20,000 USD), with demonstrated effectiveness \cite{Frey2016}. The ground was placed at the center of the forehead, approximately AFz according to the 10-20 International Standard for Electrode Placement (ISEP), and reference on the left mastoid. 6 channels were used for the 3 electrodes on each earpiece (shown in Figure \ref{fig:earpiece_diagram}); for the remaining 2 channels one AgCl ring electrode was placed on the right mastoid for later re-referencing, and one at approximately Fp1 (ISEP location above the left eye) to validate the data collected in the ears against a common scalp-based placement. Before beginning the experiment, the data from each channel was visually inspected using the OpenBCI interface. Audio stimuli were delivered through small tubes in the earpieces.

During the experiment, participants were seated in a comfortable position in a quiet room facing a laptop on which the instructions and stimuli were presented using PsychoPy \cite{peirce2007psychopy}. All tasks were performed in five trials each, followed by another set of five trials each to reduce boredom and repetition effects. Each trial was 10 seconds in length, for a total of 10 trials or 100 seconds of data collected per task. The instructions were read aloud to the participant by the experimenter, and the experiment was advanced using a pointer held in the participant's lap to minimize motion artifacts in the data. The experimenter also recorded the participant's chosen secrets for the \textit{sport}, \textit{song}, \textit{face}, \textit{speech}, and \textit{sequence} tasks and reminded the participant of these for the second set of trials. After EEG data collection, participants completed a usability questionnaire.

@@

We were able to confirm the custom-fit earpieces are able to collect quality EEG data via two metrics: good impedances measured for the ear electrodes, and alpha-band EEG activity attenuation when a participant's eyes were open versus closed.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{lrrrrrr}
& \multicolumn{6}{c}{\textbf{Impedances} [k\(\Omega\)]} \\
\cline{2-7}
& \multicolumn{3}{|c|}{\textbf{Left ear}} & \multicolumn{3}{c|}{\textbf{Right ear}} \\
\textbf{P} & \textbf{C} & \textbf{F} & \textbf{B} & \textbf{C} & \textbf{F} & \textbf{B} \\
\hline
1 & 4 & 4 & 4 & \textless1 & 4 & 3\\
2 & 9 & 5 & 4 & 3 & 4 & 4\\
3 & 4 & 5 & 4 & 9 & 6 & 9\\
4 & 4 & 5 & 4 & 3 & 16 & 9\\
5 & 9 & 20 & 7 & 3 & 7 & 9\\
6 & 5 & 8 & 2 & 1 & 1 & 9\\
7 & 2 & 9 & 8 & 7 & 5 & 6\\
\end{tabular}
\end{center}
\caption{Electrical impedances measured for concha (C), front (F) and back (B) earpiece electrodes.}
\label{tab:impedances}
\end{table}

For the alpha-attenuation test, data from the \textit{breathe} task was compared with that of the \textit{breathe - open} task. It is a well-known feature of EEG data that activity in the alpha-band (approx. 8-12 Hz) increases when the eyes are closed compared to when the eyes are open. This attenuation is clearly visible even in just a single trial's data from our ear pieces and matches that seen in our Fp1 scalp electrode data. Figure \ref{fig:alpha_atten} shows the alpha attenuation in the left ear channels, as well as Fp1 of one participant as an example. We see the same validation in the right ear channels.

@@

We analyzed the EEG signals collected during the tasks using a support vector classifier (SVC). Since past work has shown that classification tasks in EEG-based brain-computer interfaces (BCI) are linear \cite{Garrett2003a}, we used XGBoost, a popular tool for logistic linear classification \cite{Chen2016}. Compared to other linear classifiers, XGBoost uses gradient boosting; in which an algorithm generates an ensemble (in this case, a decision tree) of weak linear classifiers that minimizes a given loss function. Gradient boosting gives better results in linear classification problems, without our needing to manually tune classifier hyperparameter.

@@

For each configuration of electrodes, we calculated the mean FAR and FRR across all participants using each task as the passthought (Figure \ref{fig:meanByElectrode}). Incorporating all electrodes data results in a perfect score for all tasks. Using data from all left-ear electrodes achieves the next lowest FAR. No single electrode performs as well as the aggregate left or both-ear plus Fp1 (all) conditions.

@@

Our results indicate acceptable accuracy using data from the left ear alone. This corresponds to our desired scenario, in which the device could be worn like an earbud. As such, we focus on results from the left ear alone in our following analyses.

@@

A subset of participants completed our usability questionnaire. This questionnaire asked them to rank each of the mental tasks on 7-point likert-type scales on ease of use, level of engagement, perceived repeatability, and likeliness to use in a real-world authentication setting. The \textit{breathe} and \textit{listen} (\(\mu\)=6.75) tasks were rated easiest to use. \textit{Sequence} (\(\mu\)=5) and \textit{song} were rated highest in engagement, \textit{Breathe} (\(\mu\)=7) and \textit{listen} (\(\mu\)=6.75) were ranted highest in repeatability. In terms of likeliness of use for authentication, \textit{song-open} (\(\mu\)=5) and \textit{sequence} (\(\mu\)=4.25) were rated highest, though modestly. We also asked participants to rank the tasks from most (1) to least (9) favorite overall. The \textit{song - open} task ranked highest among favorites (\(\mu\)=4.25) followed by a tie between \textit{breathe - open}, \textit{song}, and \textit{speech} (\(\mu\)=4.75). \textit{Sequence} (\(\mu\)=7.75) and \textit{face} (\(\mu\)=6.75) were rated as least favorites.

@@

While our authentication analysis establishes that passthoughts achieve low FAR and FRR when tested against other participants' passthoughts, this does not tell us how robust passthoughts are against a spoofing attack, in which both a participant's custom-fit electrode, and details of that participant's chosen passthought, are leaked and authentication is attempted by an imposter.

The first aspect of this scenario we tested was the ability of an imposter to wear an earpiece acquired from someone else and achieve viable impedance values for EEG collection based on the fit of the pieces in their ears. P1 tried on each of the other participants' customized earpieces, which were able to at least physically fit in P1's ears. The impedances of P1 wearing other participants' earpieces were then recorded and are listed in Table \ref{tab:imposter_impedances} below. While some are a better fit than others, overall they are higher (worse) than those achieved by the pieces' intended owners themselves.

@@

To explore the scenario of an imposter attempting to gain access, we chose the case of the most vulnerable participant, P6, whose earpieces P1 had the lowest impedances while wearing. We collected data using the same data collection protocol, but had P1 refer to P6's report of chosen passsthoughts.
P1 performed each of P6's passthoughts (simulating an "inside imposter" from within the system). Following the same classification and preprocessing steps, we generated 200 samples per task for our imposters, using data from all left ear electrodes.

@@

Since every participant has one classifier per task (for which that task is the passthought), we are able to make 200 spoofed attempts with the correct passthought on each of P6's classifiers. We find zero successful spoof attempts for tasks with a chosen secret (e.g., \textit{song} or \textit{face}). However, we also do not find any successful spoof attacks for tasks with no chosen secret (e.g., \textit{breathe}). In fact, in all 1,800 spoof attempts (200 attempts for each of the 9 classifiers), we do not find a single successful attack on any of P6's classifiers.

However, since this participant's data appeared in the initial pool, the classifier may have been trained on his or her recordings as negative examples. To explore the efficacy of an outsider spoofing recordings, we repeated the same protocol with an individual who did not appear in our initial set of participants (an "outside imposter"). Again, we find zero successful authentications out of 1,800 attempts.

@@

Our findings demonstrate the apparent feasibility of a passthoughts system consisting of a single earpiece with three electrodes and a reference, all on or around the left ear. FARs and FRRs are very low across all participants and tasks, with FARs overall lower than FRRs, a desirable pattern in terms of authenticating access to potentially sensitive information. Participants' best-performing passthoughts typically see no errors in our training. Furthermore, no spoofed attacks were successful in our cursory analysis.

This study was conducted with a small, relatively homogeneous sample of participants. In order to establish the validity of this system for widespread real-world use we feel it is necessary to expand the size and diversity of participants in future studies. In the case of this initial evaluation however, the homogeneity of our participant pool strengthens the reported results given that system is meant to distinguish between individuals. 
%We feel that a greater number of  participants from a more diverse pool would greatly improve our understanding and interpretation of the capabilities of this technology. 
%This sample size is in line with this area of research as it is comparable to that of other scalp-based EEG passthoughts work\cite{Ashby2011, Marcel2007a, Palaniappan2008, palaniappan2006electroencephalogram, Poulos2002, Chuang2013b}, the only other in-ear EEG passthoughts study we are aware of \cite{curran2016passthoughts} (notably a generic fit system was used here), and other custom-fit in-ear EEG research \cite{Kidmose2013a}. 
This sample size is comparable to that of other EEG passthoughts work \cite{Ashby2011, Marcel2007a, Poulos2002, Chuang2013b, curran2016passthoughts} and other custom-fit in-ear EEG research \cite{Kidmose2013a, Mikkelsen2015}. 
The fitting and manufacturing of custom-fit earpieces was the main limitation to increasing our sample size, and may very well pose a limitation in proliferation and adoption of such a technology. Recently, there have been developments in at-home kits for creating one's own custom-fitted earpieces \cite{voix2015settable} that could help overcome this barrier.
